Empirical Evaluation of eval in R

### 2020-08-18 inst/analysis/analysis_stable.Rmd
Classification of language expr (!st argument of eval)

@@

Expressions of types other than `call`,  `expression`, `name`, `bytecode` and `promise` are just returned unchanged. Here `language` refers to `call` , `name`to `symbol`. `bytecode` refers to bytecode objects, that are generated by the `compile`function:

@@

>>>>

We look at the first argument of `eval`, the `expr` argument. It is the expression to be evaluated. It can be of different types, mainly usual numeric types, vectors, strings, and language types: calls, symbols, promise, bytecode and expressions.

`bytecode` refers to bytecode objects, that are generated by the `compile`function:

Type `expression` is a sequence of calls and symbols.

@@

Expressions of types other than `call`,  `expression`, `name`, `bytecode` and `promise` are just returned unchanged.

@@

### `expr` argument when it is a call

```{r expr_call_arg}
expr_call_arg <- eval_calls %>% filter(expr_expression_type == "LANGSXP") %>% 
  mutate(function_expr = map_chr(expr_expression, function_name))
```

The following top level functions are used:

```{r top_level_function_expr}
expr_call_arg %>% count(function_expr) %>% arrange(desc(n))
```

We can distinguish a few groups:

- assignments
- reading (in a data frame, a vector)
- parsing
- defusing (`quote` and so on)
- conversion or creation of new values
- building calls and calling
- working with environments
- class assignments, reads and definitions
- building logical values
- formula
- meta programming on functions (getting access to the body of the function, to the call stack)

```{r group_expr_functions}
group_expr_functions <- expr_call_arg %>% mutate(group_function_expr = fct_collapse(function_expr, 
                          assignment = c("<-", "[[<-", "[<-", "$<-"),
                          read = c("[", "[[", "$"),
                          parsing = c("parse", "str2lang", "str2expression", "parse_only", "base::parse"),
                          defusing = c("quote", "substitute", "bquote", "expr", "quo"),
                          conversion = c("as.list", "expression", "as.name", "as.expression", "list", "as.data.frame", "as.character"),
                          calling = c("call", "as.call", "do.call"),
                          environment = c("environment", "as.environment"),
                          classAssign = c("@<-"),
                          classRead = c("slot", "@"),
                          classDef = c(".Primitive(\"class\")"),
                          logical = c("<", ">", "is.element"),
                          scope = c("{"),
                          choice = c("if"),
                          function_meta = c("sys.calls", "function", "body"),
                          formula = c("~"), # Add mathematical operators
                          other_level = "other" ))
                          #TODO: add anonymous function applications
```

```{r}
group_expr_functions %>% count(group_function_expr) %>% arrange(desc(n)) %>% knitr::kable()
```

```{r}
group_expr_functions %>% ggplot() +
  geom_bar(aes(x = fct_rev(fct_infreq(group_function_expr)))) +
  scale_y_sqrt(n.breaks = 15) +
  coord_flip() +
  labs(title = "Category of calls", y = "n", x = "category" ) 
```
